---
title: "AIDS Drug"
author: "Carla Seals"
date: "May 18, 2019"
output: 
  html_document:
    number_section: true
    toc: true
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
avast <-read.csv(file="avastivironal.csv",header=TRUE)
avastIV = filter(avast, Ivdrug=='user')
avastM = filter(avast, sex=='M')
avastF = filter(avast, sex=='F')
avastLGBT = filter(avast, LBGT=='yes')
avastNLGBT = filter(avast, LBGT=='no')
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative AIDS (avastivirinol) drug is under test in clinical trials. We will analyze the effectiveness of this vaccination from the General population group and other subgroups such as by gender(male and female), IV drug users, and by sexual orientation. 

##Question 1


Did the vaccine make a difference in the general population?  


$H_0:$ The vaccination had no effect of the contracting of AIDS.

$H_A:$ The vaccination changed the possibility of contracting AIDS.


##Question 2

Does the variable of gender change the effectiveness of the vaccination?  



$H_0:$ No difference in effectiveness between males and females.


$H_A:$ Effectiveness of the vaccination changes between males and females.


##Question 3


Does the variable of using drugs change the effectiveness of the vaccination?


$H_0:$ No difference in effectiveness between IV users and IV non-users.


$H_A:$ Effectiveness of the drug changes between users and non-users.

##Question 4

Does the variable of sexual orientation change the effectiveness of the vaccination?


$H_0:$ No difference in effectiveness between LGBT and non-LGBT.


$H_A:$ Effectiveness of the drug changes between LGBT and non-LGBT.


# Methods
 <!--xtab
 chi-squared
 bar plot--!>
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

To determine the effectiveness of the vaccination, we will be doing a variable analysis with bar plots, cross tabulations, and chi-squared tests.

#Results for General effectiveness

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r, echo=FALSE}
barchartGC(~infected+treatment, data=avast, type ='percent')
```

This bar chart shows the difference between the people that contracted AIDS and those who did not contract AIDS, The effectiveness of the vaccination and the  placebo effectiveness including  the results. The group without AIDS were equally using and not using the placebo verses the vaccination. However, out of those who were infected, there was a lower percentage of people who recieved the vaccination. We can determine that the vaccination prevented a considerable amount of people from contracting AIDS.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r, echo=FALSE}
xtabs(~treatment+infected, data = avast)
chisqtestGC(~treatment+infected, data = avast)
```

In this data we can see that between those that contracted AIDS, there is a lower amount of people infected in the general population group that recieved the vaccination than those who recieved the placebo.

The chi-square test gives us a p-value of 0 which is less than 0.05, therefore we must reject the null hypothesis in favor of the alternate hypothesis which states that the vaccinaion does effect the possiblility of contracting AIDS.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

Due to the results recieved from the bar chart, cross tabulation, and chi-squared test; we can accept the alternate hypothesis ($H_A$) that the vaccine had an influence on the general population as to whether or not they contracted aides.

#Results for Gender Effects

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
MALE ONLY CHART:
```{r, echo=FALSE}
barchartGC(~infected+treatment, data=avastM, type ='percent')
```

FEMALE ONLY CHART:
```{r, echo=FALSE}
barchartGC(~infected+treatment, data=avastF, type ='percent')
```

These charts show the differences and separate results for  the males and females effectiveness of the vaccination. The charts reveals that the female community is negatively effected by the vaccination whereas the male community is positively effected. The women are more likely to get AIDS than the men.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
MALES:
```{r,echo=FALSE}
xtabs(~treatment+infected, data = avastM)
chisqtestGC(~treatment+infected, data = avastM)
```
FEMALES:
```{r, echo=FALSE}
xtabs(~treatment+infected, data = avastF)
chisqtestGC(~treatment+infected, data = avastF)
```

Cross tabulations for both male and female groups show differences in effectiveness of the vaccination.

Both chi-square tests also reveal a p-value that is less than 0.05 with males at 0 and females at 1e-04. this proves that sex is a factor in the effectiveness of the vaccination.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

According to the results of the data presented, we see that the vaccination effects men and women differently. The female population is more likely to contract AIDS when using this vaccine, while men are less likely to contract AIDS. This vaccination is giving women AIDS.

#Results for Drug Use Effects

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r, echo=FALSE}
barchartGC(~infected+treatment, data=avastIV, type ='percent')
```

This bar chart for the IV drug users show that a majority of users that got the vaccination contracted AIDS. This is the opposite of the desired effect, we want to stop people from contracting AIDS.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r, echo=FALSE}
xtabs(~treatment+infected, data = avastIV)
chisqtestGC(~treatment+infected, data = avastIV)
```


There are twice as many people who contracted AIDS than those who didn't. This shows that IV drug users is effectively negatively by this vaccination.

The chi-square test shows a p-value of 0 which is less than 0.05. Therefore we reject the null hypothesis and accept the alternate hypothesis that IV drug users change the effectiveness of the vaccination.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

From the results seen here, the vaccination has a negative effect on those who are IV drug users. Those who use and get the vaccine are more likely to contract AIDS than those who do not use IV drugs.

#Results for Sexual Orientation Effects

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r, echo=FALSE}
barchartGC(~infected+treatment, data=avastLGBT, type ='percent')
```

This table shows that the drug has a positive effect on those of the LGBT community. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r, echo=FALSE}
xtabs(~treatment+infected, data = avastLGBT)
chisqtestGC(~treatment+infected, data = avastLGBT)
```
The numerical summary show the  cross tabulation that the vaccinated people in the LGBT community are three times less likely to contract AIDS.


The chi-square test shows a p-value of 0 which is less than 0.05. Therefore we reject the null hypothesis and accept that the vaccination  has an effect on the likelyhood of contracting AIDS.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

According to the bar charts, chi-squared test and cross tabulations performed in this group, sexual orientation does not change the effectiveness of the vaccination.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Overall, the community is positively effected by the vaccination. However, the vaccination has a negative effect on both females and IV drug users. The variable of sexual orientation has little to no effect on the outcome of the vaccination effectiveness.

In only one case we accept the null hypothesis. This is the hypothesis that sexual orientation has no effect on the effectiveness of the experimental vaccination.
